---
title: "Practical Machine Learning/ Prediction Assignment"
author: "ieta hamdan"
date: "Monday, December 23, 2015"
output: html_document
---
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement Â- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this data set, the participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants toto predict the manner in which praticipants did the exercise.

The dependent variable or response is the "classe" variable in the training set.



```{r}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(2048)
options(warn=-1)
```

<h3>1)Load the data</h3>
Some values contained a "#DIV/0!" that I replaced with an NA value.

```{r}
training_data <- read.csv(file = "~/2015/051315 MSC Myprocert/Bahan Study/Module 8 Machine Learning/ASGMNT/pml-training.csv", na.strings=c("#DIV/0!") )
evaluation_data <- read.csv(file = "~/2015/051315 MSC Myprocert/Bahan Study/Module 8 Machine Learning/ASGMNT/pml-testing.csv", na.strings=c("#DIV/0!") )

```

<h3>2)Data cleaning</h3>
-Casted all columns 8 to the end to be numeric.

```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}
```

-Some columns were mostly blank.  These did not contribute well to the prediction.  I chose a feature set that only included complete columns.  I also remove user name, timestamps and windows.  

Determine and display out feature set.

```{r}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

<h3>3)Analyze the data</h3>
Create the model data built from our feature set.

```{r}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

We now build 5 random forests with 150 trees each. We make use of parallel processing to build this
model. I found several examples of how to perform parallel processing with random forests in R, this
provided a great speedup.

```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

Provide error reports for both training and test data.
```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

Conclusions and Test Data Submit
--------------------------------

As can be seen from the confusion matrix this model is very accurate.  I did experiment with PCA 
and other models, but did not get as good of accuracy. Because my test data was around 99% 
accurate I expected nearly all of the submitted test cases to be correct.  It turned out they 
were all correct.

Prepare the submission. (using COURSERA provided code)

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```
